---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load the packages for processing and plotting data.
Read the data and transform the date format into "Date" class for later analyses.
```{r, message=FALSE}
library(lubridate)
library(dplyr)
library(knitr)
library(ggplot2)
unzip(zipfile = "activity.zip")
df <- read.csv("activity.csv", header = TRUE) %>%
      tbl_df %>%
      mutate(date = ymd(date))
```


## What is mean total number of steps taken per day?
Calculate the mean and median number of steps taken per day, omitting the missing data.
```{r, results='hide'}
stepDay <- df %>%
      group_by(date) %>%
      na.omit %>%
      summarise(mean = round(mean(steps, na.rm = TRUE), 2), 
                median = median(steps, na.rm = TRUE))
```


1. Make a histogram of the total number of steps taken each day
```{r}
res <- hist(stepDay$mean, plot = FALSE)
lim <- round(range(res$counts), -1)
hist(stepDay$mean, xlab = "", main = "Total Number of Steps Taken Each Day", ylim = lim)  ## get the range of counts by 
```


2. Calculate and report the **mean** and **median** total number of steps taken per day
```{r}
kable(stepDay, format = "html")
```


## What is the average daily activity pattern?
Calcuate the average number of steps taken by intervals.
```{r, results='hide'}
times <- df %>%
      group_by(interval) %>%
      summarise(ave = round(mean(steps, na.rm = TRUE), 2))
```


1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
plot(times$ave ~ times$interval, type = "l", xlab = "Interval", ylab = "Average Number of Steps Taken")
```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
times[times$ave == max(times$ave),]$interval
```


## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)
```{r}
sum(is.na(df))
```


2. Filling in the missing values in the dataset by using the mean for that 5-minute interval.

```{r}
repna.df <- df %>%
      group_by(interval) %>%
      mutate(steps = ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps))
```


3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
newDf <- repna.df %>%
      ungroup 
newstepDay <- newDf %>%
      group_by(date) %>%
      summarise(mean = round(mean(steps), 2) , median = round(median(steps), 2))
```


4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. 
```{r}
hist(newstepDay$mean, xlab = "", main = "Total Number of Steps Taken Each Day", sub = "Imputing Missing Data")
```


```{r}
kable(newstepDay, format = "html")
```


##### The new table shows a different picture from the previous estimation, especially the median values from imputing values are higher than other observations.

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
weekend <- c("Saturday", "Sunday")
weekDf <- newDf %>%
      mutate(day = factor(weekdays(date) %in% weekend, labels = c("weekday", "weekend")))
```

2. Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using **simulated data**:
```{r}
weeks <- weekDf %>%
      group_by(day, interval) %>%
      summarise(ave = round(mean(steps, na.rm = TRUE), 2))

ggplot(data = weeks, aes(interval, ave)) +
      geom_line() +
      facet_grid(day ~ .) +
      xlab("Interval") +
      ylab("Average Number of Steps Taken") 
```